---
title: "campaign_food_spending"
author: "Dylan Manfre"
date: "2023-08-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#libraries
library(tidyverse)
library(tidycensus)
library(janitor)
library(lubridate)
#install.packages("ipumsr")
library(ipumsr)
```

*** Early notes from Aug. 23 ***

- Years filtered for: 2011 to today
- States filtered fsor: DC, MD, VA
- Search terms I used in the FEC.gov site.
  - food, drinks, meals, catering, dining, beverage, meeting, events (produced a dataset with 274,352 results)
  - Link: https://www.fec.gov/data/disbursements/?data_type=processed&two_year_transaction_period=2012&two_year_transaction_period=2014&two_year_transaction_period=2016&two_year_transaction_period=2018&two_year_transaction_period=2020&two_year_transaction_period=2022&two_year_transaction_period=2024&min_date=01%2F01%2F2011&max_date=12%2F31%2F2024&recipient_state=DC&recipient_state=MD&recipient_state=VA&disbursement_description=FOOD&disbursement_description=DRINKS&disbursement_description=MEALS&disbursement_description=CATERING&disbursement_description=DINING&disbursement_description=BEVERAGE&disbursement_description=MEETING&disbursement_description=EVENTS
  

  

- We want to identify the purpose category as "Fundraising" or another term and make sure it doesn't say "travel" so we exclude those unnecessary transactions.



```{r}

# loading in the campaign_food_data.csv
campaign_food_data <- read_csv("data/campaign_food_data.csv")
head(campaign_food_data)

```


** I put in 7 filters to the FEC site with the above link. Shouldn't then there only be 7 categories that show up? **

```{r}
#counts for categories
campaign_food_data %>% 
group_by(disbursement_description) %>% 
count() %>% 
arrange(desc(n))

#recipient totals
category_totals <- campaign_food_data %>% 
group_by(recipient_name, recipient_state) %>% 
summarize(total_spent = sum(disbursement_amount)) %>% 
arrange(desc(total_spent))

#state totals
state_totals <-  campaign_food_data %>% 
group_by(recipient_state) %>% 
summarize(total_spent = sum(disbursement_amount)) %>% 
arrange(desc(total_spent))
#why is DC not showing up?
#Look for the NAs in the dataset

#use this block to try and filter for the NAs
campaign_food_data <- campaign_food_data %>% 
filter(!is.na(disbursement_amount))
#omitted image_number: 12020611921 becasue it showed up as a NA and FEC site said it was $0 anyway.
```


```{r}
#trying to see how many transactions there are in the years, based on the quater of the year?

#Then try to find out how much money was spent in that quarter
#Try to set up the columns: YEAR, QUARTER, NO. OF TRANSACTIONS, Q_TOTAL

#Which year had the most transactions
year_totals <-campaign_food_data %>% 
group_by(report_year) %>% 
count() %>%
arrange(desc(n))
#2018 had most transactions

#Which Quarter of 2018 had most
2018_quarters <-  campaign_food_data %>%
  filter(report_year == "2018", 
         report_type %in% c("Q1", "Q2", "Q3", "Q4"))
#Create a month and year column from the date column. 






```

** Should I use Open Refine to standardize some of the disbursement_descriptions
List of ones to standardist: FOOD AND BEVERAGE, FOOD & BEVERAGE, FOOD/BEVERAGES, FOOD / BEVERAGES, FOOD AND MEALS, FOOD & BEV, FOOD AND BEVERAGE EXPENSE

** RESTURANT MEALS IS A GOOD ONE TO HONE IN ON **


```{r}




```






